---
title: "Athletes_R"
output: html_notebook
---

```{r}
df <- read.csv('athletes_events.csv')
head(df)
```

### 1. How old were the youngest male and female participants of the 1996 Olympics?

for females
```{r}
min(df[df['Sex'] == 'F' & df['Year'] == 1996, 'Age'], na.rm=T)
```
for males
```{r}
min(df[df['Sex'] == 'M' & df['Year'] == 1996, 'Age'], na.rm=T)
```

### 2. What was the percentage of male gymnasts among all the male participants of the 2000 Olympics? Round the answer to the first decimal.

```{r}
df1 <- df[df['Sport'] == 'Gymnastics' & df['Year'] == 2000 & df['Sex'] == 'M',]
df2 <- df[df['Year'] == 2000 & df['Sex'] == 'M',]
round(length(unique(df1$Name))/length(unique(df2$Name))*100, 1)
```

### 3. What are the mean and standard deviation of height for female basketball players participated in the 2000 Olympics? Round the answer to the first decimal.

mean
```{r}
df3 <- df[df['Sport'] == 'Basketball' & df['Sex'] == 'F' & df['Year'] == 2000,]
round(mean(df3$Height), 1)
```

std
```{r}
round(sd(df3$Height), 1)
```

### 4. Find a sportsperson participated in the 2002 Olympics, with the highest weight among other participants of the same Olympics. What sport did he or she do?

```{r}
df[df$Year == 2002 & !is.na(df$Weight) & df$Weight == max(df[df$Year == 2002,]['Weight'], na.rm=TRUE),][1,'Sport']
```

### 5. How many times did Pawe Abratkiewicz participate in the Olympics held in different years?

```{r}
df4 = df[df$Name == 'Pawe Abratkiewicz',]
nrow(df4[!duplicated(df4$Games),])
```

### 6. How many silver medals in tennis did Australia win at the 2000 Olympics?

```{r}
nrow(df[df['Year'] == 2000 & df['Team'] == 'Australia' & df['Sport'] == 'Tennis' & !is.na(df$Medal) & df['Medal'] == 'Silver',])
```

### 7. Is it true that Switzerland won fewer medals than Serbia at the 2016 Olympics? Do not consider NaN values in Medal column.

```{r}
nrow(df[df['Year'] == 2016 & df['Team'] == 'Switzerland' & !is.na(df$Medal),]) < nrow(df[df['Year'] == 2016 & df['Team'] == 'Serbia' & !is.na(df$Medal),])
```

### 8. What age category did the fewest and the most participants of the 2014 Olympics belong to?

```{r}
pm = df[df['Year'] == 2014 & (55 >= df['Age']) & (df['Age'] >= 45), ]
mid = df[df['Year'] == 2014 & (45 > df['Age']) & (df['Age'] >= 35), ]
ya = df[df['Year'] == 2014 & (35 > df['Age']) & (df['Age'] >= 25), ]
tn = df[df['Year'] == 2014 & (25 > df['Age']) & (df['Age'] >= 15), ]

count <- c(length(unique(pm$Name)), length(unique(mid$Name)), length(unique(ya$Name)), length(unique(tn$Name)))
categories <- c('45-55', '35-45', '25-35', '15-25')
df5 <- data.frame(count, categories)
```
the most
```{r}
df5[df5$count == max(df5$count), 'categories']
```
the fewest
```{r}
df5[df5$count == min(df5$count), 'categories']
```

### 9. Is it true that there were Summer Olympics held in Lake Placid? Is it true that there were Winter Olympics held in Sankt Moritz?

for Lake Placid
```{r}
print(nrow(df[(df['City'] == 'Lake Placid') & (df$Season == 'Summer'),]) > 0)
```
for Sankt Moritz
```{r}
print(nrow(df[(df['City'] == 'Sankt Moritz') & (df$Season == 'Winter'),]) > 0)
```

### 10. What is the absolute difference between the number of unique sports at the 1996 Olympics and 2016 Olympics?

```{r}
df6 = df[df$Year == 1996,]
df7 = df[df$Year == 2016,]
abs(length(unique(df6$Sport)) - length(unique(df7$Sport)))
```
