---
title: "R Notebook"
output: html_notebook
---

```{r}
install.packages("plotly")
```

```{r}
library(dplyr)
```


```{r}
library(plotly)
```

```{r}
df <- read.csv('2008.csv.bz2')
```

```{r}
head(df)
```

```{r}
dim(df)
```

```{r}
colnames(df)
```

# The tasks

### 1. Find top-10 carriers in terms of the number of completed flights (UniqueCarrier column)?
#### Which of the listed below is not in your top-10 list?

* DL
* AA
* OO
* EV

Answer: EV

```{r}
barplot(sort(table(df['UniqueCarrier']), decreasing = TRUE)[0:10])
```

### 2. Plot distributions of flight cancellation reasons (CancellationCode).
#### What is the most frequent reason for flight cancellation?

* carrier (A)
* weather conditions (B)
* National Air System (C)
* security reasons (D)

Answer: weather conditions

```{r}
barplot(sort(table(df['CancellationCode']), decreasing=TRUE)[2:5])
```

### 3. Which route is the most frequent, in terms of the number of flights?
#### (Take a look at 'Origin' and 'Dest' features. Consider A->B and B->A directions as different routes)

* New-York – Washington
* San-Francisco – Los-Angeles
* San-Jose – Dallas
* New-York – San-Francisco

Answer: San-Francisco – Los-Angeles

```{r}
df2 <- df %>% group_by(Origin, Dest) %>% summarise(count = n()) %>% arrange(desc(count))
df2[df2$count == max(df2$count),]
```

### 4. Find top-5 delayed routes (count how many times they were delayed on departure). From all flights on these 5 routes, count all flights with weather conditions contributing to a delay.

* 449
* 539
* 549
* 668

Answer: 668

```{r}
filter(df, DepDelay>0) %>% group_by(Origin, Dest) %>% summarize(count = n(), w = length(which(WeatherDelay > 0))) %>% arrange(desc(count)) %>% .[1:5, 'w'] %>% sum()
```

### 5. Examine the hourly distribution of departure times. For that, create a new series from DepTime, removing missing values.
#### Choose all correct statements:

* Flights are normally distributed within time interval [0-23] (Search for: Normal distribution, bell curve).
* Flights are uniformly distributed within time interval [0-23].
* In the period from 0 am to 4 am there are considerably less flights than from 7 pm to 8 pm.

Answer: In the period from 0 am to 4 am there are considerably less flights than from 7 pm to 8 pm.

```{r}
df3 <- df[!is.na(df['DepTime']), c('DepTime','Year')]
df3$DepTime <- sapply(df3$DepTime, function(x) (x %/% 100) %% 24)
barplot(table(df3['DepTime']))
```

### 6. Show how the number of flights changes through time (on the daily/weekly/monthly basis) and interpret the findings.
#### Choose all correct statements:

* The number of flights during weekends is less than during weekdays (working days).
* The lowest number of flights is on Sunday.
* There are less flights during winter than during summer.

Answer: The number of flights during weekends is less than during weekdays (working days) AND There are less flights during winter than during summer.

```{r}
barplot(table(df['Month']))
```

```{r}
barplot(table(df['DayOfWeek']))
```

```{r}
table(df['Month'])
table(df['Month'])[1] + table(df['Month'])[2] + table(df['Month'])[12] < table(df['Month'])[6] + table(df['Month'])[7] + table(df['Month'])[8]
```

### 7. Examine the distribution of cancellation reasons with time. Make a bar plot of cancellation reasons aggregated by months.
#### Choose all correct statements:

* December has the highest rate of cancellations due to weather.
* The highest rate of cancellations in September is due to Security reasons.
* April's top cancellation reason is carriers.
* Flights cancellations due to National Air System are more frequent than those due to carriers.

Answer: December has the highest rate of cancellations due to weather AND April's top cancellation reason is carriers.

```{r}
df4 <- df[df$Cancelled == 1, c('CancellationCode','Month')]
df5 <- df4 %>% group_by(CancellationCode, Month) %>% count(CancellationCode)
df5
barplot(formula = df5$n ~ df5$CancellationCode + df5$Month)
#black is 'A' (Carrier)
#grey is 'B' (Weather)
#light-grey is 'C' (National Air System)
```

### 8. Which month has the greatest number of cancellations due to Carrier?

* May
* January
* September
* April

Answer: April

```{r}
df6 <- filter(df, CancellationCode == 'A')
barplot(table(df6['Month']))
```

### 9. Identify the carrier with the greatest number of cancellations due to carrier in the corresponding month from the previous question.

* 9E
* EV
* HA
* AA

Answer: AA

```{r}
sort(table(df6['UniqueCarrier']), decreasing = TRUE)[1]
```

### 10. Examine median arrival and departure delays (in time) by carrier. Which carrier has the lowest median delay time for both arrivals and departures? Leave only non-negative values of delay times ('ArrDelay', 'DepDelay'). Boxplots can be helpful in this exercise, as well as it might be a good idea to remove outliers in order to build nice graphs. You can exclude delay time values higher than a corresponding .95 percentile.

* EV
* OO
* AA
* AQ

```{r}

df %>% filter(ArrDelay > 0) %>% group_by(UniqueCarrier) %>% summarize(n = median(ArrDelay)) %>% arrange(n) %>% .[1, ]
```

```{r}
df %>% filter(DepDelay > 0) %>% group_by(UniqueCarrier) %>% summarize(n = median(DepDelay)) %>% arrange(n) %>% .[1, ]
```

```{r}
arr_outliers <- quantile(df$DepDelay, 0.95, na.rm = TRUE)
arr_outliers
df7 <- df[!is.na(df$ArrDelay) & 0 < df$ArrDelay & df$ArrDelay <= arr_outliers, c('ArrDelay', 'UniqueCarrier')]
```

```{r}
dep_outliers <- quantile(df$DepDelay, 0.95, na.rm = TRUE)
dep_outliers
df8 <- df[!is.na(df$DepDelay) & 0 < df$DepDelay & df$DepDelay <= dep_outliers, c('DepDelay', 'UniqueCarrier')]
```

```{r}
boxplot(df7$ArrDelay~df7$UniqueCarrier, xlab="Carriers", ylab="Delays",
        col=c("aliceblue", "azure4", "orange", "salmon", "purple", "chocolate4", "gray14", "royalblue1", "peachpuff", "red", "green",
              "cyan4", "antiquewhite", "yellow", "mediumblue", "navajowhite4", "gray32", "burlywood", "turquoise1", "palegreen1"))
```

```{r}
table(df['UniqueCarrier'])
```

```{r}
boxplot(df8$DepDelay~df8$UniqueCarrier, xlab="Carriers", ylab="Delays",
        col=c("aliceblue", "azure4", "orange", "salmon", "purple", "chocolate4", "gray14", "royalblue1", "peachpuff", "red", "green",
              "cyan4", "antiquewhite", "yellow", "mediumblue", "navajowhite4", "gray32", "burlywood", "turquoise1", "palegreen1"))
```